---
title: "Store Sales Forecasting"
subtitle: "Kaggle Competition"
author: "Mateo"
format: 
  html:
    toc: true
    embed-resources: true
---


# Initial setup

## Packages
```{r}
#install.packages("devtools")
#install.packages("tictoc")
#install.packages("future")
#install.packages("parallelly")

```

## Libraries
```{r}
#| label: pkgs
#| message: false
#| warning: false

library(conflicted)
## install kaggler package from github
# pak::pak("mkearney/kaggler")
library(future)
library(furrr)
library(kaggler)
library(tidymodels)
library(tidyverse)
library(tsibble)
library(feasts)
library(fable)
library(fabletools)
library(purrr)
library(dplyr)

conflicts_prefer(
  lubridate::date(),
  dplyr::filter()
)
```

### Usamos multisession:
```{r}
options(future.globals.maxSize = 2 * 1024^3)  # 2 GB
plan(multisession, workers = 4)  # Usa varios núcleos
```

# Data import

## `train.csv`

### Convert to a `tsibble`

```{r}
#| label: train
train_tsbl <- read_csv("data/train.csv", 
                       show_col_types = FALSE) |> 
  as_tsibble(index = date, key = c(store_nbr, family)) |> 
  select(-onpromotion)

train_tsbl
```

### Checking for gaps

```{r}
train_tsbl |> 
  has_gaps()
```
All of them have implicit gaps in time. Below we'll find where are such gaps:

```{r}
gaps <- train_tsbl |> 
  count_gaps(.full = TRUE) |> 
  distinct(.from)

gaps
```

It's Christmas. We'll fill the implicit gaps and set to zero its value.

```{r}
#| label: fill_gaps
train_tsbl <- train_tsbl |> 
  fill_gaps(.full = TRUE, 
            sales = 0L)

train_tsbl
```
## additional data

### `holidays_events.csv`

```{r}
holidays <- read_csv("data/holidays_events.csv", show_col_types = FALSE)
holidays
```

### `stores.csv`

```{r}
stores <- read_csv("data/stores.csv", show_col_types = FALSE)
stores
```

### `oil.csv`

```{r}
oil <- read_csv("data/oil.csv", show_col_types = FALSE) |> 
  as_tsibble(index = date) |> 
  fill_gaps(.full = TRUE) |> 
  fill(dcoilwtico, .direction = "downup")
oil 
```

### Joining the data in one `tsibble`

```{r}
train_tsbl <- train_tsbl |> 
  left_join(oil, by = "date")
train_tsbl
```

### `test.csv`

```{r}
test <- read_csv("data/test.csv", show_col_types = FALSE) |> 
  select(-onpromotion)
test
```


# Exploratory analysis

## Products with no sales

We'll check if there are any family products not being sold in specific stores:

```{r}
ceros <- train_tsbl |> 
  as_tibble() |> 
  group_by(store_nbr, family) |> 
  summarise(sales = sum(sales), .groups = "drop") |> 
  filter(sales == 0) |> 
  select(-sales)

ceros
```
There are `{r} nrow(ceros)` series with no sales in the training set. We'll use a **NAIVE** model to forecast them.The time series that do have sales will be stored in `series_tsbl`:

```{r}
series_tsbl <- train_tsbl |> 
  anti_join(ceros, by = c("store_nbr", "family"))

series_tsbl
```

and the ones without sales in `series_ceros`:

```{r}
series_ceros <- train_tsbl |> 
  anti_join(series_tsbl, by = c("date", "store_nbr", "family"))
series_ceros
```

The forecasts for such series are done below:

```{r}
series_ceros_fit <- series_ceros |> 
  model(naive = NAIVE(sales))
series_ceros_fit

series_ceros_fcst <- series_ceros_fit |> 
  forecast(h = 16)

series_ceros_fcst
```

First, several questions arise about our data, two in particular:
1. Will there be time series with many zeros at the beginning?
2. Will there be time series with only zeros at the end?

## We filter out the zeros

### First question:

```{r}
# We look for the first non-null value in each series:
primer_valor_no_cero <- series_tsbl |> 
  group_by(store_nbr, family) |>  
  filter(sales != 0) |>  
  slice_min(date) |>  # We choose the first row (the first non-zero value)
  select(store_nbr, family, date) |> 
  arrange(desc(date)) |> 
  rename(fecha_inicio = date)
primer_valor_no_cero
```

```{r}
# Filter the series by removing data prior to the first non-zero value
ceros_series_filtradas <- series_tsbl |> 
  left_join(primer_valor_no_cero, by = c("store_nbr", "family")) |>  
  filter(date < fecha_inicio) |>  # We perform the join only when the date is less than 'fecha_inicio'
  select(-fecha_inicio)  # We remove the 'fecha_inicio' column from the resulting table

# Filtramos y guardamos las series importantes
series_tsbl_filtered <- series_tsbl |> 
  anti_join(ceros_series_filtradas, by = c("date", "store_nbr", "family"))

series_tsbl_filtered
```

Now for the second question, simply filter out those time series where the last 30 days had sales = 0.
This makes your Forecast equal to 0.

### Second question:
```{r}
# Last month:
series_tsbl_ultimo_mes <- series_tsbl_filtered |>
  filter(date >= as.Date("2017-07-15"), date <= as.Date("2017-08-15"))

# Let's see which ones are all zero. 
# I'll also take this opportunity to exclude from my forecasts those that had very few sales.
series_pocas_ventas_ultimo_mes <- series_tsbl_ultimo_mes |>
  as_tibble() |> 
  group_by(store_nbr, family) |> 
  summarise(sales = sum(sales), .groups = "drop") |> 
  filter(sales < 4) |> 
  select(-sales)
series_pocas_ventas_ultimo_mes
```

So let's make their forecast:
```{r}
# First we will remove the data from these 82 time series from the important tsibble:
series_tsbl_importantes <- series_tsbl_filtered |> 
  anti_join(series_pocas_ventas_ultimo_mes, by = c("store_nbr", "family"))

# Now we work with 'series_pocas_ventas_ultimo_mes'
series_pobres_ultimo_mes <- series_pocas_ventas_ultimo_mes |>
  left_join(series_tsbl_ultimo_mes, by = c("store_nbr", "family")) |>
  as_tsibble(index = date, key = c(store_nbr, family))

# Now the forecast:
series_pobres_ultimo_mes_fit <- series_pobres_ultimo_mes |> 
  model(naive = NAIVE(sales))

series_pobres_ultimo_mes_fcst <- series_pobres_ultimo_mes_fit |> 
  forecast(h = 16)

series_pobres_ultimo_mes_fcst
```

## Simple models
Now that I've filtered some time series based on some conditions related to business rules.
I can filter out those time series that predict well using a simple model.
```{r}

```




















# Benchmark model

## Fit

```{r}
tictoc::tic()
series_tsbl_fit <- series_tsbl_importantes |> 
  model(
    ets = ETS(sales)
  )
tictoc::toc()

series_tsbl_fit
```

```{r}
tictoc::tic()
fabletools::accuracy(series_tsbl_fit)
tictoc::toc()
```


## Forecast

```{r}
tictoc::tic()
series_tsbl_fcst <- series_tsbl_fit |> 
  forecast(h = 16)
tictoc::toc()

series_tsbl_fcst
```

Joining forecasts

```{r}
full_fcst <- series_tsbl_fcst |> 
  bind_rows(series_ceros_fcst) |>
  bind_rows(series_pobres_ultimo_mes_fcst)

full_fcst
```


## Submission

```{r}
submission <- full_fcst |> 
  left_join(test, by = c("date", "store_nbr", "family")) |> 
  as_tibble() |>
  select(id, .mean) |>
  rename(sales = .mean)

submission
```


```{r}
submission |> 
  write_csv("VersionesPrueba/submission_M3.csv")
```

## we finished using parallelism

```{r}
plan(sequential)
```